---
title: "STA 3100 Programming with Data: Assignment 030"
params:
   term: "Fall 2020"
   duedate: "2020-11-24"
   section: "3502"
   classnumber: "27256"
   name: "Regine Louis"
   ufid: "70592702"
   email: "rlouis1@ufl.edu"
   total_points: 40
   grade: NA
output: 
   html_document: default
   pdf_document: default
---

# Working with the New York Times COVID-19 Data

As we have seen in class, COVID-19 data collected by the *New York
Times* are available in a repository on Github.  In this assignment we
will work with these data to reproduce some of the elements of the
*Times*'s reporting on COVID-19.  

# Exercises

1. The file `us.csv` contains aggregated data for the entire U.S.  In
   this file, the `cases` and `deaths` columns represent *cumulative*
   cases and deaths due to covid. The `diff()` function can be used to
   compute the differences between each consecutive element of a
   vector, so it could be used to compute the daily numbers of cases
   and deaths.  However, `diff()` returns a vector of length one less
   than the length of the original vector (e.g., diff(c(1, 3, 6, 10))
   returns 2, 3, 4) and this can make it somewhat
   inconvenient to use when transforming columns of data frames.
   
   An alternative is to us the more general `filter()` function with
   an appropriate choice of the `filter` and `sides` arguments.  Thus function
   can also be used to computing running averages and similar quantities.
   
    a. Read the file `us.csv` into R as the data frame `us` and do the
       following:
	  
        * Transform the `date` column into a column of class `Date`.
        * Use `filter()` to add a column named `new_cases` containing the
          number of new cases reported on each date. The first
          value in this column will be `NA`.
        * Use `filter()` to add a column named `new_deaths` containing the
          number of new deaths reported on each date. The first
          value in this column will be `NA`.
        * Use `filter()` to add a column named `avg_new_cases` where
     	    each element represents the mean number of new cases for the
     	    previous 7 days (inclusive of the current day).  The first 7
     	    values in this column will be `NA`.
        * Use `filter()` to add a column named `avg_new_deaths` where
     	    each element represents the mean number of new deaths for the
     	    previous 7 days (inclusive of the current day).  The first 7
     	    values in this column will be `NA`.
```{r}

us <- read.csv(file = 'us.csv', header=TRUE)

#changing date column to a column of class#
us$date<- as.Date(us$date)
#check to see if it works# 
str(us)

#adding new cases column #
new_cases<- stats::filter(us$cases, filter= c(1, -1), method = c("convolution"), sides = 1) 
us$new_cases <- new_cases

#new deaths#
new_deaths<- stats::filter(us$deaths, filter= c(1, -1), method = c("convolution"), sides = 1) 
us$new_deaths <- new_deaths

#average new cases#
avg_new_cases<- stats::filter(us$new_cases, filter= c(1/7, 1/7, 1/7, 1/7, 1/7, 1/7, 1/7), method = c("convolution"), sides = 1) 
us$avg_new_cases <-avg_new_cases

#avg new deaths#
avg_new_deaths<- stats::filter(us$new_deaths, filter= c(1/7, 1/7, 1/7, 1/7, 1/7, 1/7, 1/7), method = c("convolution"), sides = 1) 
us$avg_new_deaths <-avg_new_deaths



```
       Note that the `filter()` function used here is
       `stats::filter()` from the `stats` package, which is loaded by
       default in R. (The `dplyr` package has a completely different
       `filter()` function which plays an important role in the
       "tidyverse".  If you have problems using `filter()`, you should
       make sure that you do NOT have the `dplyr` package loaded.  If
       you do, then you will need to explicitly type out
       `stats::filter()` to get the `stats` version.)
     	  
        
     	  
    b. Create a plot of daily cases similar to the one found at the
  	   top of [this
  	   page](https://www.nytimes.com/interactive/2020/us/coronavirus-us-cases.html).
  	   Plot only data beginning from 2020-03-01. (Note that this plot
  	   and a similar plot for deaths appear again about 1/3 of the way
  	   down the page.)

       Try to do this using the formula method of the `plot()` function with the
       optional arguments `type = "h"`, `col = "gray"`, and `data = us`
       and using the `subset` argument to plot only the data for dates 2020-03-01
       and after. (You may also wish to experiment with the optional argument `lwd`.)
       
       Then use the formula interface to the `lines()` function to add the curve
       showing the seven-day running average. (Again, you may wish to experiment
       with the optional argument `lwd`.)

        

    c. Repeat part (b) for deaths.

``` {r}

#plot of daily cases#
plot(new_cases~ date, type="h", col= "gray", main= "COVID-19 Cases in the US", xlab= "Date", ylab= "Cases", data = us, subset= (us$date >= "2020-03-01"))
lines(us$avg_new_cases~date, col="black", data = us, subset= (us$date>= "2020-03-01"))


#plot of daily death#
plot(new_deaths~ date, type="h", col= "gray",main= "COVID-19 Deaths in the US", xlab= "Date", ylab= "Deaths", data = us, subset= (us$date >= "2020-03-01"))
lines(us$avg_new_deaths~date, col="black", data= us, subset=(us$date>= "2020-03-01"))

```

2. The file `us-states.csv` contains county-level data for the U.S.

    a. Read `us-states.csv` into R as the data frame `us_states`
       and transform the date column into a column of class `Date`.
	   
        

    b. Use `subset()` to extract the data for the state of Florida
       and save it as a data frame named `florida`. Be sure that the
	   rows are correctly ordered by date, and then repeat parts 1b
	   and 1c of this assignment for Florida, i.e., plot the number of
	   daily new cases and deaths, along with their 7-day running averages.

        
```  {r}

#States Data#
us_states <- read.csv(file = 'us-states.csv', header=TRUE)

#changing date column to a column of class#
us_states$date<- as.Date(us_states$date)
#check to see if it works# 
str(us_states)

#subsetting florida #
florida <- subset (us_states, date>= "2020-03-01" & state == "Florida", select= c("date", "state", "fips", "cases", "deaths"))

#Florida's new cases/deaths/ avg 7 days#
#adding new cases column #
new_fl_cases<- stats::filter(florida$cases, filter= c(1, -1), method = c("convolution"), sides = 1) 
florida$new_fl_cases <- new_fl_cases

#new deaths#
new_fl_deaths<- stats::filter(florida$deaths, filter= c(1, -1), method = c("convolution"), sides = 1) 
florida$new_fl_deaths <- new_fl_deaths

#average new cases#
avg_new_fl_cases<- stats::filter(florida$new_fl_cases, filter= c(1/7, 1/7, 1/7, 1/7, 1/7, 1/7, 1/7), method = c("convolution"), sides = 1) 
florida$avg_new_fl_cases <-avg_new_fl_cases

#avg new deaths#
avg_new_fl_deaths<- stats::filter(florida$new_fl_deaths, filter= c(1/7, 1/7, 1/7, 1/7, 1/7, 1/7, 1/7), method = c("convolution"), sides = 1) 
florida$avg_new_fl_deaths <-avg_new_fl_deaths

#plot of daily cases#
plot(new_fl_cases~ date, type="h", col= "gray", main= "COVID-19 Cases in Florida", xlab= "Date", ylab= "Cases", data = florida, subset= (florida$date >= "2020-03-01"))
lines(florida$avg_new_fl_cases~date, col="black", data = florida, subset= (florida$date>= "2020-03-01"))

#plot of daily death#
plot(new_fl_deaths~ date, type="h", col= "gray",main= "COVID-19 Deaths in Florida", xlab= "Date", ylab= "Deaths", data = florida, subset= (florida$date >= "2020-03-01"))
lines(florida$avg_new_fl_deaths~date, col="black", data= florida, subset=(florida$date>= "2020-03-01"))


```
        

        

3. The file `us-counties.csv` contains county-level data for the U.S.

    a. Read `us-counties.csv` into R as the data frame `us_counties`
       and transform the date column into a column of class `Date`.
	   
        

    b. Use `subset()` to extract the data for Alachua County, Florida,
       and save it as a data frame named `alachua`. Be sure that the
	   rows are correctly ordered by date, and then repeat part 1b
	   this assignment for Alachua County, i.e., plot the number of daily new
	   cases along with their 7-day running average. 

```{r}
#County Data#
us_counties <- read.csv(file = 'us-counties.csv', header=TRUE)
#changing date column to a column of class#
us_counties$date<- as.Date(us_counties$date)
#check to see if it works# 
str(us_counties)

#subset for Alachua, Florida#
alachua <- subset (us_counties, date>= "2020-03-01" & state == "Florida" & county == "Alachua", select= c("date", "state","county", "fips", "cases", "deaths"))

#Alachua's new cases/deaths/ avg 7 days#
#adding new cases column #
new_al_cases<- stats::filter(alachua$cases, filter= c(1, -1), method = c("convolution"), sides = 1) 
alachua$new_al_cases <- new_al_cases

#new deaths#
new_al_deaths<- stats::filter(alachua$deaths, filter= c(1, -1), method = c("convolution"), sides = 1) 
alachua$new_al_deaths <- new_al_deaths

#average new cases#
avg_new_al_cases<- stats::filter(alachua$new_al_cases, filter= c(1/7, 1/7, 1/7, 1/7, 1/7, 1/7, 1/7), method = c("convolution"), sides = 1) 
alachua$avg_new_al_cases <-avg_new_al_cases

#avg new deaths#
avg_new_al_deaths<- stats::filter(alachua$new_al_deaths, filter= c(1/7, 1/7, 1/7, 1/7, 1/7, 1/7, 1/7), method = c("convolution"), sides = 1) 
alachua$avg_new_al_deaths <-avg_new_al_deaths

#plot of daily cases#
plot(new_al_cases~ date, type="h", col= "gray", main= "COVID-19 Cases in Alachua County, Florida", xlab= "Date", ylab= "Cases", data = alachua, subset= (alachua$date >= "2020-03-01"))
lines(alachua$avg_new_al_cases~date, col="black", data = alachua, subset= (alachua$date>= "2020-03-01"))

#plot of daily death#
plot(new_al_deaths~ date, type="h", col= "gray",main= "COVID-19 Deaths in Alachua County, Florida", xlab= "Date", ylab= "Deaths", data = alachua, subset= (alachua$date >= "2020-03-01"))
lines(alachua$avg_new_al_deaths~date, col="black", data= alachua, subset=(florida$date>= "2020-03-01"))



```
	   Q: What do you notice when comparing the plot of daily new cases
       in Alachua county to the analogous plot for the state of
	   Florida as a whole?  What might explain what you are seeing?
```{r}
# In the state of Florida, record number of COVID-19 cases peaked in the months of July/August. In Alachua county, cases were realitvely stable during July and August, however, in the months of September/Octover cases reached an all time high. The differences observed between Florida COVID-19 cases and Alachua county covid cases can be explained by the large number of college students returning to Alachua county for school. #


```

        

        
